---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library("pomp")
library("ggplot2")
```



*This is the example I tried to modify:* \newline https://kingaa.github.io/sbied/stochsim/notes.pdf \newline
*There are some exercises, solutions can be found here:* \newline https://kingaa.github.io/sbied/stochsim/exercises.html \newline 
*Similar example modified here:* \newline
http://kingaa.github.io/short-course/stochsim/stochsim.html \newline
`load("worked_example_data.RData")`

```{r warnings = F}
# loading observed data
library(tidyverse)
read_csv(paste0("https://kingaa.github.io/sbied/stochsim/", "Measles_Consett_1948.csv")) %>% select(week,reports = cases) -> meas
meas %>% as.data.frame() %>% print.data.frame()
```

### SIR model
```{r}
sir_step <- function (S, I, R, N, Beta, mu_IR, delta.t, ...) {
  dN_SI <- rbinom(n=1,size=S,prob=1-exp(-Beta*I/N*delta.t))
  dN_IR <- rbinom(n=1,size=I,prob=1-exp(-mu_IR*delta.t))
  S <- S - dN_SI
  I <- I + dN_SI - dN_IR
  R <- R + dN_IR
  c(S = S, I = I, R = R) # Return S I R at the end of each step
}
```

```{r}
# initialise the model using N (population size) and eta, some fraction of susceptibles 
sir_rinit <- function (N, eta, ...) {
  c(S = round(N * eta),
    I = 1,
    R = round(N * (1 - eta)))
}
```

```{r}
library(pomp)
meas %>%
  pomp(
    times = "week",
    t0 = 0,
    rprocess = euler(sir_step, delta.t = 1 / 7),
    rinit = sir_rinit
  ) -> measSIR  
```

### SIRH

`H` is an accumulator variable. It is reset every week and "observes" the unreported cases (or so I understand). I don't quite understand where/why we would introduce this in our (Tom's) model. Besides, the code won't run with H, so I've completely excluded it. Which then messes up the meaning of k and p (see dmeas and rmeas). 

```{r}
sir_step <- function (S, I, R, H, N, Beta, mu_IR, delta.t, ...) {
  dN_SI <- rbinom(n=1,size=S,prob=1-exp(-Beta*I/N*delta.t))
  dN_IR <- rbinom(n=1,size=I,prob=1-exp(-mu_IR*delta.t))
  S <- S - dN_SI
  I <- I + dN_SI - dN_IR
  R <- R + dN_IR
  H <- H + dN_IR;
c(S = S, I = I, R = R, H = H)
}

sir_rinit <- function (N, eta, ...) {
  c(
    S = round(N * eta),
    I = 1,
    R = round(N * (1 - eta)),
    H = 0
  )
}
```

```{r}
measSIR %>%
  pomp(
    rprocess = euler(sir_step, delta.t = 1 / 7),
    rinit = sir_rinit,
    accumvars = "H"
  ) -> measSIR
```

### dmeas and rmeas functions

```{r}
# dmeas will be used for likelihood estimation 
sir_dmeas <- function(reports, H, rho, k, log, ...) {
  dnbinom(
    x = reports,
    size = k,
    mu = rho * H,
    log = log
  )
}
# rmeas will be used for simulation
# effectively, probability of a case (ie the data) ~ NegBin (pH, k)
# where p = probability of reporting/diagnosing/observing a case
# H = true incidence, an accumulator variable which is reset after every observation(?)
sir_rmeas <- function (H, rho, k, ...) {
  c(reports = rnbinom(n = 1, size = k, mu = rho * H))
}
```

```{r}
measSIR %>%
  pomp(rmeasure = sir_rmeas,
       dmeasure = sir_dmeas) -> measSIR
```


```{r}
## Running some simulations and plotting the output. 
measSIR %>%
  simulate(params = c(Beta = 30, mu_EI = 0.8, mu_IR = 1.3, 
                      rho = 0.5, k = 10, eta = 0.06, N = 38000),nsim = 20,
           format = "data.frame", include.data = TRUE) %>%
  ggplot(aes(x = week, y = reports, group = .id, color = .id == "data")) +
  geom_line() + guides(color = "none") + theme_bw()
```

C snippets which achieve the exact same end. Not used bc Li said Csnippets are impossible to debug. 

```{r}
seir_step <- Csnippet("
  double dN_SE = rbinom(S,1-exp(-Beta*I/N*dt));
  double dN_EI = rbinom(E,1-exp(-mu_EI*dt));
  double dN_IR = rbinom(I,1-exp(-mu_IR*dt));
  S -= dN_SE;
  E += dN_SE - dN_EI;
  I += dN_EI - dN_IR;
  R += dN_IR;
  H += dN_IR;
")


seir_init <- Csnippet("
  S = nearbyint(eta*N);
  E = 0;
  I = 1;
  R = nearbyint((1-eta)*N);
  H = 0;
")

measSIR %>%
  pomp(
    rprocess=euler(seir_step,delta.t=1/7),
    rinit=seir_init,
    paramnames=c("N","Beta","mu_EI","mu_IR","rho","eta"),
    statenames=c("S","E","I","R","H")
  ) -> measSEIR
```













