---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library("pomp")
library("ggplot2")
library("ukcovid19")
```



# PHE data 

The following chunks of code retrieve data about daily COVID-19 cases and deaths up to the date of latest publication. To retreive data used in code below use `load("pomp_phe_data.RData")`. 

```{r eval = T, include = T}
library("ukcovid19")

query_filters <- c(
    'areaType=nation'
)

cases_and_deaths = list(
     date = "date",
     areaName = "areaName",
     areaCode = "areaCode",
     newCasesByPublishDate = "newCasesByPublishDate",
     cumCasesByPublishDate = "cumCasesByPublishDate",
     newDeaths28DaysByPublishDate = "newDeaths28DaysByPublishDate",
     cumDeaths28DaysByPublishDate = "cumDeaths28DaysByPublishDate"
)

data <- get_data(
    filters = query_filters, 
    structure = cases_and_deaths
)

data$date = as.Date(data$date)
```

```{r eval = T, include = T}
# Filtering cases between Oct 2020 and Feb 2021  
# filter(date >= "2021-05-01")
covid = data %>% dplyr::filter(date <= "2021-02-01") %>%
    dplyr::filter(date >= "2020-10-01")
covid = dplyr::select(covid, newCasesByPublishDate, date)
```

```{r}
ggplot(covid, aes(x = date, y = newCasesByPublishDate)) +
    geom_line() + theme_classic() +
    xlab("") + scale_x_date(date_labels = "%m-%Y") +
    ylab("New Daily Cases")
```

```{r}
# df for pomp
covid2 = data.frame(cases = covid$newCasesByPublishDate, 
                    time = seq_along(covid$newCasesByPublishDate))
```


## Tom's waning immunity code 

Kindly borrowed as a guide for modifying the example given in "worked_example.RMD". 

```{r eval = F, include = T}
### Exponential times
S = S + omega * R - beta * S * I
E = E + beta * S * I - E * sigma
I = I + E * sigma - I * gamma
R = R + I * gamma - omega * R

### Gamma-distributed times
S[(i + 1)] = S[i] + omega * dt * R2[i] * o - beta * dt * S[i] * (I1[i] + I2[i])
E1[(i + 1)] = E1[i] + beta * dt * S[i] * (I1[i] + I2[i]) - E1[i] * sigma * dt * m
E2[(i + 1)] = E2[i] + sigma * dt * m * (E1[i] - E2[i])
E3[(i + 1)] = E3[i] + sigma * dt * m * (E2[i] - E3[i])
E4[(i + 1)] = E4[i] + sigma * dt * m * (E3[i] - E4[i])
I1[(i + 1)] = I1[i] + E4[i] * sigma * dt * m - I1[i] * gamma * dt * n
I2[(i + 1)] = I2[i] + gamma * dt * n * (I1[i] - I2[i])
R1[(i + 1)] = R1[i] + I2[i] * gamma * dt * n - omega * dt * R1[i] * o
R2[(i + 1)] = R2[i] + omega * dt * o * (R1[i] - R2[i])
```


# SEIR model 

The simple model below runs, albeit giving odd results. \newline
`load("working_pomp_code.RData")` \newline

```{r}
# change during each time step 
SEIR_step <- function(S, E, I, R, N, Beta, sigma, gamma, Omega, delta.t, ...) {
    dN_RS = rbinom(1, R, 1 - exp(-Omega*(R/N)*delta.t)) #i'm not sure if this is correct...
    dN_SE = rbinom(1, S, 1 - exp(-Beta*((S*I)/N)*delta.t))
    dN_EI = rbinom(1, E, 1 - exp(-(E/N)*sigma*delta.t))
    dN_IR = rbinom(1, I, 1 - exp(-(I/N)*gamma*delta.t))
    S = S + dN_RS - dN_SE
    E = E + dN_SE - dN_EI
    I = I + dN_EI - dN_IR
    R = R + dN_IR - dN_RS
    return(c(S = S, E = E, I = I, R = R))
} 

# iniital conditions
SEIR_step_rinit <- function(N, eta,...) {
    c(S = round(N*eta), E = 0, I = 1, R = round(N*(1 - eta)))
}

# parameters to be used in simulation 
paramts = c(Beta = 0.91205, sigma = 0.2222, gamma = 0.3257, Omega = 0.01111, 
            p = 0.5, k = 10, N = 66435550, eta = 0.999)
#Beta=30,mu_EI=0.8,mu_IR=1.3,p=0.5,k=10,eta=0.06,N=38000)
#paramts = c(Beta = 0.91205, sigma = 0.2222, gamma = 0.3257, omega = 0.01111)

# dmeas process used for parameter estimation (likelohood) 
SEIR_dmeas <- function(cases, p, k, log, ...) {
    dnbinom(x = cases, size = k, mu = p, log = log)
}

# rmeas process used for simulation 
SEIR_rmeas <- function(p, k, ...) {
    c(cases = rnbinom(n = 1, size = k, mu = (p)))
} # mu = p*H    I havw removed H as a state from the model which modifies the meaning of the rmeas process.... but I am not sure if what I've done makes sense 

# coding the model as a pomp object 
covidSEIR <- pomp(data = covid2,
                  times = "time",t0 = 0,
                  rprocess = euler(SEIR_step, delta.t = 0.1),
                  rinit = SEIR_step_rinit,
                  rmeasure = SEIR_rmeas,
                  dmeasure = SEIR_dmeas,
                  paramnames = c("Beta", "sigma", "gamma", "Omega",
                                 "p", "k", "N", "eta"),
                  statenames = c("S", "E", "I", "R")
)
```

```{r}
simSEIR = covidSEIR %>% 
  simulate(params = c(Beta = 0.91205, sigma = 0.2222, 
                      gamma = 0.3257, Omega = 0.01111,
                      p = 0.75, k = 10, N = 66435550, eta = 0.999),
    nsim = 20, format = "data.frame", include.data = T)
```

```{r}
ggplot(simSEIR, aes(x = time, y = cases, group = .id, color = .id == "data")) + 
  geom_line() + guides(color = "none") + theme_bw()
# the results do not make sense. could be attributed to poor parameter and initial condition choice. could the flow of recovered into susceptible be problematic?  
```

```{r}
str(simSEIR)
length(covid2$cases)
## interestingly , all the data in covid2$cases is also present in simSEIR$cases 
length(which(covid2$cases %in% simSEIR$cases)) == length(covid2$cases)
identical(which(covid2$cases %in% simSEIR$cases), seq_along(covid2$cases))
## but there are 10k observations in simSEIR$cases, 
##  we have simulated SOMETHING albeit once 
length(simSEIR$cases)
```






## Next steps: attempting to adapt gamma-distributed compartment times

The following code does not work but would have been a good next step in adapting the model. 

```{r eval = F, include = T}
SEIR_step <- function(S, E, I, R, N, Beta, sigma, gamma, Omega, delta.t, ...) {
  SE1 = rbinom(1, S, 1 - exp(-beta*dt*S*(I1 + I2)))
  E1I = rbinom(1, E1, 1 - exp(-E1*sigma*dt*m))
  E12 = rbinom(1, E1, 1 - exp(-sigma*dt*m*(E1 - E2)))
  E23 = rbinom(1, E2, 1 - exp(-sigma*dt*m*(E2 - E3)))
  E34 = rbinom(1, E3, 1 - exp(-sigma*dt*m*(E3 - E4)))
  E4I1 = rbinom(1, E4, 1 - exp(-E4*sigma*dt*m))
  I1R = rbinom(1, I1, 1 - exp(-I1*gamma*dt*n))
  I12 = rbinom(1, I1, 1 - exp(-gamma*dt*n*(I1 - I2)))
  I2R1 = rbinom(1, I2, 1 - exp(-I2*gamma*dt*n))
  R1S = rbinom(1, R1, 1 - exp(-omega*dt*R1*o))
  R12 = rbinom(1, R1, 1 - exp(-omega*dt*o*(R1 - R2)))
  R2S = rbinom(1, R2, 1 - exp(-omega*dt*R2*o))
  S = S + R2S - SE1
  E1 = E1 + SE1 - E1I1
  E2 = E2 + E12
  E3 = E3 + E23
  E4 = E4 + E34
  I1 = I1 + E4I1 - I1R
  I2 = I2 + I12
  R1 = R1 + I2R1 - R1S
  R2 = R2 + R12
  c(S = S, E = E1+E2+E3+E4, I = I1+I2, R = R1+R2)
} 

SEIR_step_rinit <- function(N, eta,...) {
  c(S = round(N*eta), E = 0, I = 1, R = round(N*(1 - eta)))
}

#paramts = c(Beta = 0.91205, sigma = 0.2222, gamma = 0.3257, omega = 0.01111)
paramts = c(Beta = 0.91205, sigma = 0.2222, gamma = 0.3257, Omega = 0.01111, 
            p = 0.5, k = 10, N = 66435550, eta = 0.999)
#Beta=30,mu_EI=0.8,mu_IR=1.3,p=0.5,k=10,eta=0.06,N=38000),

SEIR_dmeas <- function(cases, p, k, log, ...) {
  dnbinom(x = cases, size = k, mu = p, log = log)
}

SEIR_rmeas <- function(p, k, ...) {
  c(cases = rnbinom(n = 1, size = k, mu = (p)))
}


covidSEIR <- pomp(data = covid2, 
                times = "time",t0 = 0,
                rprocess = euler(SEIR_step, delta.t = 0.1),
                rinit = SEIR_step_rinit, 
                rmeasure = SEIR_rmeas,
                dmeasure = SEIR_dmeas,
                paramnames = c("Beta", "sigma", "gamma", "Omega", 
                               "p", "k", "N", "eta"),
                statenames = c("S", "E", "I", "R")
)
```

```{r eval=F, include=F}
simSEIR = covidSEIR %>% 
  simulate(params = c(Beta = 0.91205, sigma = 0.2222, 
                      gamma = 0.3257, Omega = 0.01111,
                      p = 0.75, k = 10, N = 66435550, eta = 0.999),
    nsim = 20, format = "data.frame", include.data = T)
```

```{r eval = F, include=F}
ggplot(simSEIR, aes(x = time, y = cases, group = .id, color = .id == "data")) + 
  geom_line() +
  guides(color = "none") + theme_bw()
```


